Carbonate chemistry at Ny-Ålesund

### 2021-03-19 awipev-CO2.Rmd
Calculate spec pH fb; kf="dg" everywhere

# Conversion of spec pH of in situ samples to in situ temp and 4 °C
pH_seafet <- d %>%
  dplyr::filter(!is.na(ph_s_sf) & !is.na(temp_insitu_ctd))
pH_seafet <- pH_seafet %>% 
  dplyr::mutate(
    ph_s_sf_temp_insi = pHinsi(pH = pH_seafet$ph_s_sf, 
                         ALK = pH_seafet$at_mean_interp*1e-6, Tinsi = pH_seafet$temp_insitu_ctd, 
                         Tlab = pH_seafet$temp_s_lab, Pinsi = pH_seafet$pressure_insitu_ctd/10, 
                         S = pH_seafet$sal_mix, Pt = 0, Sit = 0, k1k2 = "l", kf = "x", ks="d", 
                         pHscale = "T", b="u74"),
   ph_s_sf_norm4 = pHinsi(pH = pH_seafet$ph_s_sf, 
                         ALK = pH_seafet$at_mean_interp*1e-6, Tinsi = 4, 
                         Tlab = pH_seafet$temp_s_lab, Pinsi = pH_seafet$pressure_insitu_ctd/10, 
                         S = pH_seafet$sal_mix, Pt = 0, Sit = 0, k1k2 = "l", kf = "x", ks="d", 
                         pHscale = "T", b="u74")
  )

# Conversion of spec pH of FerryBox samples to FerryBox temp
pH_durafet <- d %>%
  dplyr::filter(!is.na(ph_s_dur) & !is.na(sal_fb) & !is.na(temp_fb))
pH_durafet <- pH_durafet %>% 
  dplyr::mutate(ph_s_dur_temp_fb = pHinsi(pH = pH_durafet$ph_s_dur, ALK = pH_durafet$at_mean_interp*1e-6, Tinsi = pH_durafet$temp_fb, Tlab = pH_durafet$temp_s_lab, Pinsi = 0, S = pH_durafet$sal_fb, Pt = 0, Sit = 0))

@@

# d_all <- d_all %>%
#   dplyr::mutate(ph_s_sf_norm4 = ifelse(!is.na(pH_seafet$ph_s_sf),
#                                               pHinsi(pH = pH_seafet$ph_s_sf, 
#                          ALK = pH_seafet$at_mean_interp*1e-6, Tinsi = 4, 
#                          Tlab = pH_seafet$temp_s_lab, Pinsi = pH_seafet$pressure_insitu_ctd/10, 
#                          S = pH_seafet$sal_mix, Pt = 0, Sit = 0, k1k2 = "l", kf = "x", ks="d", 
#                          pHscale = "T", b="u74"), NA
#   ))

@@

pH_sf$pHint_tot_sf_norm4 <- pHinsi(pH=pH_sf$pHint_tot_sf, ALK=mean(d_all$at, na.rm=TRUE), Tinsi=4, Tlab=pH_sf$temp_insitu_ctd_9m, Pinsi=11/10, S=pH_sf$sal_insitu_ctd_9m, Pt=0, Sit=0, k1k2 = "l", kf = "x", ks="d", pHscale = "T", b="u74")

pH_dur <- d_all %>%
  dplyr::select(datetime, sal_fb,temp_fb, ph_dur_corr)%>%
  dplyr::filter(!is.na(ph_dur_corr), !is.na(temp_fb), !is.na(sal_fb))
pH_dur$pH_durafet_norm4 <- pHinsi(pH=pH_dur$ph_dur_corr, ALK=mean(d_all$at, na.rm=TRUE), Tinsi=4, Tlab=pH_dur$temp_fb, Pinsi=11/10, S=pH_dur$sal_fb, Pt=0, Sit=0, k1k2 = "l", kf = "x", ks="d", pHscale = "T", b="u74")

@@

### **Calculated / Durafet pHs**
```{r pHs comparaison dur calc, echo=FALSE, message= FALSE, warning=FALSE}
# Normalize phcalc from AT/CT to 4°c
pH_calc <- discrete %>%
  dplyr::select(datetime, sal_fb, temp_fb, phcalc_fb, at)%>%
  dplyr::filter(!is.na(phcalc_fb), !is.na(temp_fb), !is.na(sal_fb))
pH_calc$pH_calc_fb_norm4 <- pHinsi(pH = pH_calc$phcalc_fb, ALK=pH_calc$at*1e-6, Tinsi=4, Tlab=pH_calc$temp_fb, Pinsi=11/10, S=pH_calc$sal_fb, Pt=0, Sit=0, k1k2 = "l", kf = "x", ks="d", pHscale = "T", b="u74")

@@

## Impact of choice of constants  {.tabset .tabset-pills}
### Lueker vs Sulpis using means of monthly means
```{r impact of constants 1, echo=FALSE, warning=FALSE, message=FALSE}
# Compare Lueker and Sulpis
# use mean values calculated for impacts of nutrients above

# first, order columns
d_all <- d_all %>% 
  dplyr::select(order(colnames(d_all))) %>% 
  dplyr::select(datetime, date, year, monthd, montht, everything())

# # calculate means
# monthly_means <- d_all %>% 
#   group_by(montht) %>% 
#   summarize(mean_ct = mean(ct, na.rm = TRUE),
#             mean_at_calc = mean(at_calc, na.rm = TRUE),
#             mean_pco2_calc = mean(pco2_calc_fb, na.rm = TRUE),
#             mean_sal = mean(sal_fb, na.rm = TRUE),
#             mean_temp = mean(temp_fb, na.rm = TRUE)) %>% 
#   ungroup()
# means <- monthly_means %>% 
#   summarise(across(
#     .cols = is.numeric, 
#     .fns = mean, na.rm = TRUE, 
#     .names = "{col}"
#   ))

mean_ct <-  mean(discrete$ct, na.rm = TRUE)
mean_at <-  mean(discrete$ct, na.rm = TRUE)
mean_pco2 <-  mean(d_all$pco2_calc_fb)
mean_sal <-  mean(discrete$sal_fb, na.rm = TRUE)
mean_temp <-  mean(discrete$temp_fb, na.rm = TRUE)

# calculate typical uncertainty assuming uncertainties of 3 uatm for pCO2 and 4 umol/kg for AT
errors_pco2_at <- seacarb::errors(flag=24, var1=mean_pco2_calc, var2=1e-6*mean_at_calc, S=mean_sal,
                          T=mean_temp, P=0, Pt=0, Sit=0, k1k2="l", kf="dg", ks="d", pHscale="T",
                          b="u74", evar1=3, evar2 = 4e-6)
errors_at_ct <- seacarb::errors(flag=15, var1=1e-6*mean_at_calc, var2=1e-6*mean_ct, S=mean_sal,
                          T=mean_temp, P=0, Pt=0, Sit=0, k1k2="l", kf="dg", ks="d", pHscale="T",
                          b="u74", evar1=4e-6, evar2 = 4e-6)
errors <- tibble(e_pH = c(errors_pco2_at$pH, errors_at_ct$pH),
                 e_pCO2 = c(NA, errors_at_ct$pCO2),
                 e_DIC = c(errors_pco2_at$pH, NA)*1e3,
                 e_Oa = c(errors_pco2_at$OmegaAragonite, errors_at_ct$OmegaAragonite)
)

# impacts using pco2 and AT pair
carb_l_1 <- carb(24, mean_pco2_calc, mean_at_calc*1e-6, S = mean_sal, T = mean_temp, P = 0, 
                 Pt = 0, Sit = 0,
                 k1k2 = "l", kf = "dg", ks = "d", pHscale = "T", b = "u74")
carb_s_1 <- carb(24, mean_pco2_calc, mean_at_calc*1e-6, S = mean_sal, T = mean_temp, P = 0, 
                 Pt = 0, Sit = 0,
                 k1k2 = "s20", kf = "dg", ks = "d", pHscale = "T", b = "u74")
diff_ls_1 <- carb_l_1 - carb_s_1

# impacts using AT-DIC pair
carb_l_2 <- carb(15, mean_at_calc*1e-6, mean_ct*1e-6, S = mean_sal, T = mean_temp, P = 0, 
                 Pt = 0, Sit = 0,
                 k1k2 = "l", kf = "dg", ks = "d", pHscale = "T", b = "u74")
carb_s_2 <- carb(15, mean_at_calc*1e-6, mean_ct*1e-6, S = mean_sal, T = mean_temp, P = 0,
                Pt = 0, Sit = 0, 
                k1k2 = "s20", kf = "dg", ks = "d", pHscale = "T", b = "u74")
diff_ls_2 <- carb_l_2 - carb_s_2
const_choice <- bind_rows(diff_ls_1, diff_ls_2) %>% 
  dplyr::mutate(flag = c("pCO2 and ALK", "ALK and DIC"),
                S = mean_sal,
                T = mean_temp,
                DIC = DIC * 1e6
                ) %>% 
  dplyr::select(flag, pH, pCO2, DIC, OmegaAragonite) %>% 
  dplyr::bind_cols(errors)
```
The means used are:
`r knitr::kable(means, format = "simple")`

@@

### Lueker vs Sulpis using all values pCO2 and at_calc
```{r impact of constants 2, echo=FALSE, warning=FALSE, message=FALSE}
# Compare Lueker and Sulpis using pco2_fb and at_calc
tmp <- d_all %>% 
  tidyr::drop_na(pco2_fb, at_calc, sal_fb, temp_fb) %>% 
  dplyr::select(ph_s_sf_norm4, pco2_fb, at_calc, sal_fb, temp_fb)

carb_l <-
  carb(
    24,
    tmp$pco2_fb,
    tmp$at_calc * 1e-6,
    S = tmp$sal_fb,
    T = tmp$temp_fb,
    P = 0,
    Pt = 0,
    Sit = 0,
    k1k2 = "l",
    kf = "dg",
    ks = "d",
    pHscale = "T",
    b = "u74"
  )
# normalize derived pH at 4°C
carb_l$pH <- pHinsi(pH = carb_l$pH, 
                         ALK = carb_l$ALK, Tinsi = 4, 
                         Tlab = carb_l$T, Pinsi = 0, 
                         S = carb_l$S, Pt = 0, Sit = 0, k1k2 = "l", kf = "x", ks="d", 
                         pHscale = "T", b="u74")
carb_s <-
  carb(
    24,
    tmp$pco2_fb,
    tmp$at_calc * 1e-6,
    S = tmp$sal_fb,
    T = tmp$temp_fb,
    P = 0,
    Pt = 0,
    Sit = 0,
    k1k2 = "s20",
    kf = "dg",
    ks = "d",
    pHscale = "T",
    b = "u74"
  )
# normalize derived pH at 4°C
carb_s$pH <- pHinsi(pH = carb_s$pH, 
                         ALK = carb_s$ALK, Tinsi = 4, 
                         Tlab = carb_s$T, Pinsi = 0, 
                         S = carb_s$S, Pt = 0, Sit = 0, k1k2 = "l", kf = "x", ks="d", 
                         pHscale = "T", b="u74")

# do difference
diff_ls <- as_tibble(carb_l - carb_s) %>% 
  dplyr::rename(delta_ph = pH, delta_ct = DIC, delta_OmegaAragonite = OmegaAragonite) %>%
  dplyr::select(delta_ph, delta_ct, delta_OmegaAragonite)

# calculate descriptive statistics
diff_ls_quant <- tibble(pH = as.numeric(summary(diff_ls$delta_ph)),
                        CT = as.numeric(summary(diff_ls$delta_ct * 1e6)),
                        OmegaAragonite = as.numeric(summary(diff_ls$delta_OmegaAragonite))) %>%
  dplyr::mutate(var = c("Min", "Q1", "Median", "Mean", "Q3", "Max"), .before = 1)

# prepare tibble to compare calculated pH and spec pH
tmp2 <- tibble(ph_s_sf_norm4 = tmp$ph_s_sf_norm4, pH_l = carb_l$pH, pH_s = carb_s$pH) %>% 
  tidyr::drop_na(ph_s_sf_norm4)

# summary of differences
diff_lueker <- as.numeric(summary(tmp2$ph_s_sf_norm4 - tmp2$pH_l))
diff_sulpis <- as.numeric(summary(tmp2$ph_s_sf_norm4 - tmp2$pH_s))
lueker_sulpis_table <- tibble(var = c("Min", "Q1", "Median", "Mean", "Q3", "Max"), diff_lueker, diff_sulpis)

# make plots
p1 <- ggreg(xdata = tmp2$ph_s_sf_norm4, ydata = tmp2$pH_l, type = "MA", plot = TRUE, 
            title = "Lueker vs measured", subtitle = "both pH normalized at 4°C",
            xlab = "Spectrophotometric pH measured @ 4°C",
            ylab = "pH @ 4°C calculated (pCO2+AT) with Lueker")
p2 <- ggreg(xdata = tmp2$ph_s_sf_norm4, ydata = tmp2$pH_s, type = "MA", plot = TRUE, 
            title = "Sulpis vs measured", subtitle = "both pH normalized at 4°C",
            xlab = "Spectrophotometric pH measured @ 4°C",
            ylab = "pH @ 4°C calculated (pCO2+AT) with Sulpis") 
p <- p1 + p2
ggsave(filename = "figures/lueker-sulpis.png", plot = p, width = 20, height = 15, units = "cm")
```
The overall means of pCO2, AT (calculated from the AT-salinity relationship), salinity and temperature in the FerryBox were used to calculate pH from measured pCO2 and AT (from a salinity-AT relationship) and from AT and CT. using the formulations of Lueker et al. (2000) and Sulpis et al. (2020) of K1 and K2. The difference induced by the choice of the carbonic acid constants is large for pH, pCO2 and DIC and negligible for OmegaAragonite.

>>>>

# Conversion of spec pH of in situ samples to in situ temp and 4 °C
pH_seafet <- d %>%
  dplyr::filter(!is.na(ph_s_sf) & !is.na(temp_insitu_ctd))
pH_seafet <- pH_seafet %>% 
  dplyr::mutate(
    ph_s_sf_temp_insi = pHinsi(pH = pH_seafet$ph_s_sf, 
                         ALK = pH_seafet$at_mean_interp*1e-6, Tinsi = pH_seafet$temp_insitu_ctd, 
                         Tlab = pH_seafet$temp_s_lab, Pinsi = pH_seafet$pressure_insitu_ctd/10, 
                         S = pH_seafet$sal_mix, Pt = 0, Sit = 0, k1k2 = "l", kf = "dg", ks="d", 
                         pHscale = "T", b="u74"),
   ph_s_sf_norm4 = pHinsi(pH = pH_seafet$ph_s_sf, 
                         ALK = pH_seafet$at_mean_interp*1e-6, Tinsi = 4, 
                         Tlab = pH_seafet$temp_s_lab, Pinsi = pH_seafet$pressure_insitu_ctd/10, 
                         S = pH_seafet$sal_mix, Pt = 0, Sit = 0, k1k2 = "l", kf = "dg", ks="d", 
                         pHscale = "T", b="u74")
  )

# Conversion of spec pH of FerryBox samples to FerryBox temp
pH_durafet <- d %>%
  tidyr::drop_na(ph_s_dur, sal_fb, temp_fb)
pH_durafet <- pH_durafet %>% 
  dplyr::mutate(
    ph_s_dur_temp_fb = pHinsi(pH = pH_durafet$ph_s_dur, 
                              ALK = pH_durafet$at_mean_interp*1e-6, Tinsi = pH_durafet$temp_fb, 
                              Tlab = pH_durafet$temp_s_lab, Pinsi = 0, S = pH_durafet$sal_fb, 
                              Pt = 0, Sit = 0, k1k2 = "l", kf = "dg", ks="d", pHscale = "T", b="u74"),
   ph_s_dur_norm4 = pHinsi(pH = pH_durafet$ph_s_dur, 
                         ALK = pH_durafet$at_mean_interp*1e-6, Tinsi = 4, 
                         Tlab = pH_durafet$temp_s_lab, Pinsi = 0, 
                         S = pH_durafet$sal_fb, Pt = 0, Sit = 0, k1k2 = "l", kf = "dg", ks="d", 
                         pHscale = "T", b="u74")
   )

@@

# d_all <- d_all %>%
#   dplyr::mutate(ph_s_sf_norm4 = ifelse(!is.na(pH_seafet$ph_s_sf),
#                                               pHinsi(pH = pH_seafet$ph_s_sf, 
#                          ALK = pH_seafet$at_mean_interp*1e-6, Tinsi = 4, 
#                          Tlab = pH_seafet$temp_s_lab, Pinsi = pH_seafet$pressure_insitu_ctd/10, 
#                          S = pH_seafet$sal_mix, Pt = 0, Sit = 0, k1k2 = "l", kf = "dg", ks="d", 
#                          pHscale = "T", b="u74"), NA
#   ))

@@

pH_sf$pHint_tot_sf_norm4 <- pHinsi(pH=pH_sf$pHint_tot_sf, ALK=mean(d_all$at, na.rm=TRUE), Tinsi=4, Tlab=pH_sf$temp_insitu_ctd_9m, Pinsi=11/10, S=pH_sf$sal_insitu_ctd_9m, Pt=0, Sit=0, k1k2 = "l", kf = "dg", ks="d", pHscale = "T", b="u74")

pH_dur <- d_all %>%
  dplyr::select(datetime, sal_fb,temp_fb, ph_dur_corr)%>%
  dplyr::filter(!is.na(ph_dur_corr), !is.na(temp_fb), !is.na(sal_fb))
pH_dur$pH_durafet_norm4 <- pHinsi(pH=pH_dur$ph_dur_corr, ALK=mean(d_all$at, na.rm=TRUE), Tinsi=4, Tlab=pH_dur$temp_fb, Pinsi=11/10, S=pH_dur$sal_fb, Pt=0, Sit=0, k1k2 = "l", kf = "dg", ks="d", pHscale = "T", b="u74")

@@

### **Calculated / Durafet pHs**
```{r pHs comparaison dur calc, echo=FALSE, message= FALSE, warning=FALSE}
# Normalize phcalc from AT/CT to 4°c
pH_calc <- discrete %>%
  dplyr::select(datetime, sal_fb, temp_fb, phcalc_fb, at)%>%
  dplyr::filter(!is.na(phcalc_fb), !is.na(temp_fb), !is.na(sal_fb))
pH_calc$pH_calc_fb_norm4 <- pHinsi(pH = pH_calc$phcalc_fb, ALK=pH_calc$at*1e-6, Tinsi=4, Tlab=pH_calc$temp_fb, Pinsi=11/10, S=pH_calc$sal_fb, Pt=0, Sit=0, k1k2 = "l", kf = "dg", ks="d", pHscale = "T", b="u74")

@@

## Impact of choice of constants  {.tabset .tabset-pills}
### Lueker vs Sulpis using means of monthly means
```{r impact of constants discrete, echo=FALSE, warning=FALSE, message=FALSE}
# Compare Lueker and Sulpis using mean discrete values
# first, order columns
d_all <- d_all %>% 
  dplyr::select(order(colnames(d_all))) %>% 
  dplyr::select(datetime, date, year, monthd, montht, everything())
# calculate means of discrete values
mean_ct <- mean(discrete$ct, na.rm = TRUE)
mean_at <- mean(discrete$at, na.rm = TRUE)
mean_pco2 <- mean(discrete$pco2_calc_fb, na.rm = TRUE)
mean_sal <- mean(discrete$sal_fb, na.rm = TRUE)
mean_temp <- mean(discrete$temp_fb, na.rm = TRUE)
# calculate typical uncertainty assuming uncertainties of 3 uatm for pCO2 and 4 umol/kg for AT and CT
errors_pco2_at <- seacarb::errors(flag=24, var1=mean_pco2, var2=1e-6*mean_at, S=mean_sal,
                          T=mean_temp, P=0, Pt=0, Sit=0, k1k2="l", kf="dg", ks="d", pHscale="T",
                          b="u74", evar1=3, evar2 = 4e-6)
errors_at_ct <- seacarb::errors(flag=15, var1=1e-6*mean_at, var2=1e-6*mean_ct, S=mean_sal,
                          T=mean_temp, P=0, Pt=0, Sit=0, k1k2="l", kf="dg", ks="d", pHscale="T",
                          b="u74", evar1=4e-6, evar2 = 4e-6)
errors <- tibble(e_pH = c(errors_pco2_at$pH, errors_at_ct$pH),
                 e_pCO2 = c(NA, errors_at_ct$pCO2),
                 e_DIC = c(errors_pco2_at$pH, NA)*1e3,
                 e_Oa = c(errors_pco2_at$OmegaAragonite, errors_at_ct$OmegaAragonite)
)
# impacts using pco2 and AT pair
carb_l_1 <- carb(24, mean_pco2, mean_at*1e-6, S = mean_sal, T = mean_temp, P = 0, 
                 Pt = 0, Sit = 0,
                 k1k2 = "l", kf = "dg", ks = "d", pHscale = "T", b = "u74")
carb_s_1 <- carb(24, mean_pco2, mean_at*1e-6, S = mean_sal, T = mean_temp, P = 0, 
                 Pt = 0, Sit = 0,
                 k1k2 = "s20", kf = "dg", ks = "d", pHscale = "T", b = "u74")
diff_ls_1 <- carb_l_1 - carb_s_1
# impacts using AT-CT pair
carb_l_2 <- carb(15, mean_at*1e-6, mean_ct*1e-6, S = mean_sal, T = mean_temp, P = 0, 
                 Pt = 0, Sit = 0,
                 k1k2 = "l", kf = "dg", ks = "d", pHscale = "T", b = "u74")
carb_s_2 <- carb(15, mean_at*1e-6, mean_ct*1e-6, S = mean_sal, T = mean_temp, P = 0,
                Pt = 0, Sit = 0, 
                k1k2 = "s20", kf = "dg", ks = "d", pHscale = "T", b = "u74")
diff_ls_2 <- carb_l_2 - carb_s_2
# summary
const_choice <- bind_rows(diff_ls_1, diff_ls_2) %>% 
  dplyr::mutate(flag = c("pCO2 and ALK", "ALK and DIC"),
                S = mean_sal,
                T = mean_temp,
                DIC = DIC * 1e6
                ) %>% 
  dplyr::select(flag, pH, pCO2, DIC, OmegaAragonite) %>% 
  dplyr::bind_cols(errors)
```
The means used are:
`r knitr::kable(tibble(mean_at, mean_ct, mean_pco2, mean_sal, mean_temp), caption = "Mean discrete values.", format = "simple")`

@@

### Lueker vs Sulpis using all values pCO2 and at_calc
```{r impact of constants whole series, echo=FALSE, warning=FALSE, message=FALSE}
# Compare Lueker and Sulpis using pco2_fb and at_calc
tmp <- d_all %>% 
  tidyr::drop_na(pco2_fb, at_calc, sal_fb, temp_fb) %>% 
  dplyr::select(ph_s_sf_norm4, pco2_fb, at_calc, sal_fb, temp_fb)
# Using Lueker  
carb_l <-
  carb(
    24,
    tmp$pco2_fb,
    tmp$at_calc * 1e-6,
    S = tmp$sal_fb,
    T = tmp$temp_fb,
    P = 0,
    Pt = 0,
    Sit = 0,
    k1k2 = "l",
    kf = "dg",
    ks = "d",
    pHscale = "T",
    b = "u74"
  )
# normalize Lueker-derived pH at 4°C
carb_l$pH <- pHinsi(pH = carb_l$pH, 
                         ALK = carb_l$ALK, Tinsi = 4, 
                         Tlab = carb_l$T, Pinsi = 0, 
                         S = carb_l$S, Pt = 0, Sit = 0, k1k2 = "l", kf = "dg", ks="d", 
                         pHscale = "T", b="u74")
# Using Sulpis
carb_s <-
  carb(
    24,
    tmp$pco2_fb,
    tmp$at_calc * 1e-6,
    S = tmp$sal_fb,
    T = tmp$temp_fb,
    P = 0,
    Pt = 0,
    Sit = 0,
    k1k2 = "s20",
    kf = "dg",
    ks = "d",
    pHscale = "T",
    b = "u74"
  )
# normalize Sulpis-derived pH at 4°C
carb_s$pH <- pHinsi(pH = carb_s$pH, 
                         ALK = carb_s$ALK, Tinsi = 4, 
                         Tlab = carb_s$T, Pinsi = 0, 
                         S = carb_s$S, Pt = 0, Sit = 0, k1k2 = "s20", kf = "dg", ks="d", 
                         pHscale = "T", b="u74")
# calculate difference Lueker - Sulpis
diff_ls <- as_tibble(carb_l - carb_s) %>% 
  dplyr::rename(delta_ph = pH, delta_ct = DIC, delta_OmegaAragonite = OmegaAragonite) %>%
  dplyr::select(delta_ph, delta_ct, delta_OmegaAragonite)
# calculate descriptive statistics
diff_ls_quant <- tibble(pH = as.numeric(summary(diff_ls$delta_ph)),
                        CT = as.numeric(summary(diff_ls$delta_ct * 1e6)),
                        OmegaAragonite = as.numeric(summary(diff_ls$delta_OmegaAragonite))) %>%
  dplyr::mutate(var = c("Min", "Q1", "Median", "Mean", "Q3", "Max"), .before = 1)
# prepare tibble to compare calculated pH and spec pH
tmp2 <- tibble(ph_s_sf_norm4 = tmp$ph_s_sf_norm4, pH_l = carb_l$pH, pH_s = carb_s$pH) %>% 
  tidyr::drop_na(ph_s_sf_norm4)
# summary of differences
diff_lueker <- as.numeric(summary(tmp2$ph_s_sf_norm4 - tmp2$pH_l))
diff_sulpis <- as.numeric(summary(tmp2$ph_s_sf_norm4 - tmp2$pH_s))
lueker_sulpis_table <- tibble(var = c("Min", "Q1", "Median", "Mean", "Q3", "Max"), diff_lueker, diff_sulpis)
# make plots
p1 <- ggreg(xdata = tmp2$ph_s_sf_norm4, ydata = tmp2$pH_l, type = "MA", plot = TRUE, 
            title = "Lueker vs measured", subtitle = "both pH normalized at 4°C",
            xlab = "Spectrophotometric pH measured @ 4°C",
            ylab = "pH @ 4°C calculated (pCO2+AT) with Lueker")
p2 <- ggreg(xdata = tmp2$ph_s_sf_norm4, ydata = tmp2$pH_s, type = "MA", plot = TRUE, 
            title = "Sulpis vs measured", subtitle = "both pH normalized at 4°C",
            xlab = "Spectrophotometric pH measured @ 4°C",
            ylab = "pH @ 4°C calculated (pCO2+AT) with Sulpis") 
p <- p1 + p2
ggsave(filename = "figures/lueker-sulpis.png", plot = p, width = 20, height = 15, units = "cm")
```
The overall means of pCO2, AT (calculated from the AT-salinity relationship), salinity and temperature in the FerryBox were used to calculate pH from measured pCO2 and AT (from a salinity-AT relationship) and from AT and CT. using the formulations of Lueker et al. (2000) and Sulpis et al. (2020) of K1 and K2. The difference induced by the choice of the carbonic acid constants is large for pH, pCO2 and DIC and negligible for OmegaAragonite.
